---
title: "K43R_Fig_1 + Supplementary Fig_1"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r setup}
library(dplyr)
library(tidyverse)
library(Seurat)
library(ggplot2)
library(qs)
library(viridis)
library(patchwork)
library(data.table)
library(ggpointdensity)
library(RColorBrewer)
```


### for reduced UMAP visualization
```{r}
blank <- theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank())
```




### Load complete RRMM dataset (all patients, all time-points, BME + PCs) + metadata
```{r}
K43R_all <- qread(".../K43R_all.qs")
```


# "Classical" Seurat normalization only for this visualization, for all other analyses SCT
```{r}
K43R_all <- NormalizeData(K43R_all)
K43R_all <- FindVariableFeatures(K43R_all, selection.method = "vst", nfeatures = 2000)
K43R_all <- ScaleData(K43R_all, vars.to.regress = c("percent.mt", "G2M.Score", "S.Score"))

```


```{r fig.width=4, fig.height=3}
K43R_all <- RunPCA(object = K43R_all, npcs = 50, verbose = FALSE)
ElbowPlot(object = K43R_all, ndims = 50)
```


```{r echo=T,error = F, message = F, warning = F, verbose = F, eval=F}
K43R_all <- FindNeighbors(object = K43R_all, reduction = "pca", dims = 1:50)
K43R_all <- FindClusters(K43R_all, resolution = 0.4, verbose = F)
K43R_all <- RunUMAP(object = K43R_all, reduction = "pca", dims = 1:50, verbose = F)
```


#### Visualize
```{r fig.width=12, fig.height=4}
p1 <- DimPlot(object = K43R_all, reduction = 'umap', pt.size = 0.0001, label = F, group.by = "PID_new") + 
  ggtitle("subject") + 
  scale_color_manual(values=c(
  "dodgerblue2", 
  "#E31A1C", # red
  "green4",
  "#6A3D9A", # purple
  "maroon",
  "black", 
  "gold1",
  "skyblue2", 
  "#FB9A99", # lt pink
  "palegreen2",
  "#CAB2D6", # lt purple
  "yellow4", # lt orange
  "gray50", 
  "khaki2",
  "#FF7F00", # orange, 
  "orchid1", 
  "deeppink1", 
  "blue1", 
  "steelblue4",
  "darkturquoise", 
  "yellow4", 
  "yellow3",
  "darkorange4"
  ))+
  theme(legend.position = "none") + blank


Idents(K43R_all) <- "celltype_1_predicted"
cells.use <- names(K43R_all@active.ident[K43R_all@meta.data$celltype_1_predicted=="PC"])
p2 <- DimPlot(object = K43R_all, reduction = 'umap', pt.size = 0.5, label = F, 
              cells.highlight = cells.use, sizes.highlight=0.02, cols.highlight = "black") +
  ggtitle("SingleR prediction")  + 
  theme(legend.position = "none")+ blank


p3 <- FeaturePlot(object = K43R_all, features = "TNFRSF17", cols = c("grey", "firebrick"), pt.size =0.1, min.cutoff = "q1") + 
  theme(legend.position = "none")+ blank

px <- p1 | p2 | p3
px
```



### Cell number overview
```{r fig.width=9, fig.height=3}
df <- FetchData(K43R_all, vars = c("major_population","PID_new"))

df <- df %>% modify_if(is.character, as.factor)

df_freq <- df %>% 
  group_by(PID_new) %>% 
  dplyr::count(major_population, PID_new, .drop = F) 

p1 <- ggplot(df_freq, aes(fill=major_population, y=n, x=reorder(PID_new, -n))) + 
    geom_bar(position="dodge", stat="identity", color="black") +
  scale_fill_manual(values=c('black',"grey80"))+
 theme(axis.text.x = element_text(angle = 45, hjust = 1), panel.background=element_blank()) +
  xlab("Patient") + ylab("Number of cells") + scale_y_continuous(breaks=seq(0,20000,2000))
p1
```




### ggpointdensity split by sorting fraction
```{r fig.width=12, fig.height=4}
Idents(K43R_all) <- "sorting"
pos <- subset(K43R_all, idents = "CD138_pos")
neg <- subset(K43R_all, idents = "CD138_neg")
WBM <- subset(K43R_all, idents = "WBM")

# get embeddings
emb1 <- Embeddings(pos, reduction = "umap")
emb1 <- as.data.frame(emb1)

emb2 <- Embeddings(neg, reduction = "umap")
emb2 <- as.data.frame(emb2)

emb3 <- Embeddings(WBM, reduction = "umap")
emb3 <- as.data.frame(emb3)



p1 <- ggplot(data = emb1, mapping = aes(x=UMAP_1, y = UMAP_2)) +
  geom_pointdensity() +
  scale_color_viridis() + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  theme(legend.position = "none")+ blank

p2 <- ggplot(data = emb2, mapping = aes(x=UMAP_1, y = UMAP_2)) +
  geom_pointdensity() +
  scale_color_viridis() + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  theme(legend.position = "none")+ blank

p3 <- ggplot(data = emb3, mapping = aes(x=UMAP_1, y = UMAP_2)) +
  geom_pointdensity() +
  scale_color_viridis() + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  theme(legend.position = "none") + blank

px <- p2 | p1 | p3
px
```



### Quantify cell populations per sorting fraction
```{r fig.width=4.5, fig.height=4}
df <- FetchData(K43R_all, vars = c("sorting","major_celltype"))

df <- df %>% modify_if(is.character, as.factor)

df_freq <- df %>% 
  group_by(sorting) %>% 
  dplyr::count(major_celltype,sorting , .drop = F) %>% 
  mutate(prop=prop.table(n)) %>% 
  mutate(prop=format(prop,digits=2))
df_freq$prop <- as.numeric(as.character(df_freq$prop))

p1 <- ggplot(df_freq, aes(fill=major_celltype, y=prop, x=sorting)) + 
    geom_bar(position="dodge", stat="identity", color="black") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10), panel.background=element_blank()) +
  xlab("major cell populations") + ylab("Cell fraction") + 
  scale_fill_manual(values=c("#0072B2", "Firebrick", "#009E73","#56B4E9","#E69F00", "#F0E442"))
p1
```


### QC metrics
```{r fig.width=6, fig.height=3}
p1 <- VlnPlot(K43R_all, features = "nFeature_RNA", pt.size = 0,
              group.by = "major_celltype", sort = T, log = T, cols = c("Firebrick", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")) +
  geom_boxplot(width=.2, outlier.colour=NA, fill= "white", lwd=0.7, fatten =2) +
  theme(legend.position = "none") +
  xlab("Major_celltype") + ylab("detected genes per cell") 
p1$layers[[1]]$aes_params$size = 0.9

p2 <- VlnPlot(K43R_all, features = "nCount_RNA", pt.size = 0,
              group.by = "major_celltype", sort = T, log = T, cols = c("Firebrick", "#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442")) +
  geom_boxplot(width=.2, outlier.colour=NA, fill= "white", lwd=0.7, fatten =2) +
  theme(legend.position = "none") +
  xlab("Major_celltype") + ylab("detected UMIs per cell") 
p2$layers[[1]]$aes_params$size = 0.9

p3 <- p1 | p2
p3
```

```{r fig.width=12, fig.height=3}
Idents(K43R_all) <- "major_population"
p1 <- VlnPlot(K43R_all, features = "nFeature_RNA", pt.size = 0,
              group.by = "PID_new", idents = "PCs",
              sort = T, log = T) +
  geom_boxplot(width=.2, outlier.colour=NA, fill= "white", lwd=0.7, fatten =2) +
  theme(legend.position = "none") +
  xlab("Major_celltype") + ylab("detected genes per cell") +
  scale_fill_viridis(discrete = T)
p1$layers[[1]]$aes_params$size = 0.9


Idents(K43R_all) <- "major_population"
p2 <- VlnPlot(K43R_all, features = "nCount_RNA", pt.size = 0,
              group.by = "PID_new", idents = "PCs",
              sort = T, log = T) +
  geom_boxplot(width=.2, outlier.colour=NA, fill= "white", lwd=0.7, fatten =2) +
  theme(legend.position = "none") +
  xlab("Major_celltype") + ylab("detected UMIs per cell") +
  scale_fill_viridis(discrete = T)
p2$layers[[1]]$aes_params$size = 0.9


p3 <- p1 | p2
p3
```

```{r fig.width=12, fig.height=3}
Idents(K43R_all) <- "major_population"
p1 <- VlnPlot(K43R_all, features = "nFeature_RNA", pt.size = 0,
              group.by = "PID_new", idents = "TME",
              sort = T, log = T) +
  geom_boxplot(width=.2, outlier.colour=NA, fill= "white", lwd=0.7, fatten =2) +
  theme(legend.position = "none") +
  xlab("Major_celltype") + ylab("detected genes per cell") +
  scale_fill_viridis(discrete = T, option = "magma")
p1$layers[[1]]$aes_params$size = 0.9


Idents(K43R_all) <- "major_population"
p2 <- VlnPlot(K43R_all, features = "nCount_RNA", pt.size = 0,
              group.by = "PID_new", idents = "TME",
              sort = T, log = T) +
  geom_boxplot(width=.2, outlier.colour=NA, fill= "white", lwd=0.7, fatten =2) +
  theme(legend.position = "none") +
  xlab("Major_celltype") + ylab("detected UMIs per cell") +
  scale_fill_viridis(discrete = T, option = "magma")
p2$layers[[1]]$aes_params$size = 0.9


p3 <- p1 | p2
p3
```




# fresh/frozen controls
```{r}
Idents(K43R_all) <- "orig.ident"

#6KLSUS
P6KLSUS_WBM_fresh <- subset(K43R_all, idents = "6KLSUS_F2_H01")
P6KLSUS_WBM_fresh[["processing"]] <- "fresh"

P6KLSUS_N_fresh <- subset(K43R_all, idents = "6KLSUS_N2_H01")
P6KLSUS_N_fresh[["processing"]] <- "fresh"

P6KLSUS_T_fresh <- subset(K43R_all, idents = "6KLSUS_T2_H01")
P6KLSUS_T_fresh[["processing"]] <- "fresh"


P6KLSUS_WBM_frozen <- subset(K43R_all, idents = "6KLSUS_F2_H02")
P6KLSUS_WBM_frozen[["processing"]] <- "frozen"

P6KLSUS_N_frozen <- subset(K43R_all, idents = "6KLSUS_N2_H02")
P6KLSUS_N_frozen[["processing"]] <- "frozen"

P6KLSUS_T_frozen <- subset(K43R_all, idents = "6KLSUS_T2_H02")
P6KLSUS_T_frozen[["processing"]] <- "frozen"


#E79WZA
E79WZA_WBM_fresh <- subset(K43R_all, idents = "E79WZA_F1_H01")
E79WZA_WBM_fresh[["processing"]] <- "fresh"

E79WZA_N_fresh <- subset(K43R_all, idents = "E79WZA_N1_H01")
E79WZA_N_fresh[["processing"]] <- "fresh"

E79WZA_T_fresh <- subset(K43R_all, idents = "E79WZA_T1_H01")
E79WZA_T_fresh[["processing"]] <- "fresh"


E79WZA_WBM_frozen <- subset(K43R_all, idents = "E79WZA_F1_H02")
E79WZA_WBM_frozen[["processing"]] <- "frozen"

E79WZA_N_frozen <- subset(K43R_all, idents = "E79WZA_N1_H02")
E79WZA_N_frozen[["processing"]] <- "frozen"

E79WZA_T_frozen <- subset(K43R_all, idents = "E79WZA_T1_H02")
E79WZA_T_frozen[["processing"]] <- "frozen"


#PUVQFD
PUVQFD_WBM_fresh <- subset(K43R_all, idents = "PUVQFD_F1_H01")
PUVQFD_WBM_fresh[["processing"]] <- "fresh"

PUVQFD_N_fresh <- subset(K43R_all, idents = "PUVQFD_N1_H01")
PUVQFD_N_fresh[["processing"]] <- "fresh"

PUVQFD_T_fresh <- subset(K43R_all, idents = "PUVQFD_T1_H01")
PUVQFD_T_fresh[["processing"]] <- "fresh"


PUVQFD_WBM_frozen <- subset(K43R_all, idents = "PUVQFD_F1_H02")
PUVQFD_WBM_frozen[["processing"]] <- "frozen"

PUVQFD_N_frozen <- subset(K43R_all, idents = "PUVQFD_N1_H02")
PUVQFD_N_frozen[["processing"]] <- "frozen"

PUVQFD_T_frozen <- subset(K43R_all, idents = "PUVQFD_T1_H02")
PUVQFD_T_frozen[["processing"]] <- "frozen"


PUVQFD <- merge(PUVQFD_WBM_fresh, list(PUVQFD_N_fresh, PUVQFD_T_fresh, PUVQFD_WBM_frozen, PUVQFD_N_frozen, PUVQFD_T_frozen))
E79WZA <- merge(E79WZA_WBM_fresh, list(E79WZA_N_fresh, E79WZA_T_fresh, E79WZA_WBM_frozen, E79WZA_N_frozen, E79WZA_T_frozen))
P6KLSUS <- merge(P6KLSUS_WBM_fresh, list(P6KLSUS_N_fresh, P6KLSUS_T_fresh, P6KLSUS_WBM_frozen, P6KLSUS_N_frozen, P6KLSUS_T_frozen))

MM_all <- merge(PUVQFD, list(E79WZA, P6KLSUS))

```


### Plot QC metrics
```{r fig.width=8, fig.height=2.5}
df <- FetchData(MM_all, vars = c("nFeature_RNA", "PID_new", "processing", "percent.mt", "nCount_RNA"))

dodge <- position_dodge(width = 0.8)
p1 <- ggplot(df, aes(y=nFeature_RNA, x=PID_new)) + 
    geom_violin(aes(fill=processing),position=dodge, lwd =0.5) +
    geom_boxplot(aes(group=interaction(processing,PID_new)), 
            width=0.25, fill="white", position=dodge,
            outlier.shape=NA) +
 theme(axis.text.x = element_text(angle = 45, hjust = 1),
       panel.background=element_blank(), 
       axis.line = element_line(size = 0.5, colour = "black"), legend.position = "none")+
  scale_y_continuous(trans = 'log10')

dodge <- position_dodge(width = 0.8)
p2 <- ggplot(df, aes(y=nCount_RNA, x=PID_new)) + 
    geom_violin(aes(fill=processing),position=dodge, lwd =0.5) +
    geom_boxplot(aes(group=interaction(processing,PID_new)), 
            width=0.25, fill="white", position=dodge,
            outlier.shape=NA) +
 theme(axis.text.x = element_text(angle = 45, hjust = 1),
       panel.background=element_blank(), 
       axis.line = element_line(size = 0.5, colour = "black"), legend.position = "none") +
  scale_y_continuous(trans = 'log10')

dodge <- position_dodge(width = 0.8)
p3 <- ggplot(df, aes(y=percent.mt, x=PID_new)) + 
    geom_violin(aes(fill=processing),position=dodge, lwd =0.5) +
    geom_boxplot(aes(group=interaction(processing,PID_new)), 
            width=0.25, fill="white", position=dodge,
            outlier.shape=NA) +
 theme(axis.text.x = element_text(angle = 45, hjust = 1),
       panel.background=element_blank(), 
       axis.line = element_line(size = 0.5, colour = "black"), legend.position = "none")

p4 <- p1 | p2| p3
p4
```




```{r}
blas_set_num_threads(1)
omp_set_num_threads(1)
options(future.globals.maxSize=50*1024^3) 
PUVQFD <- SCTransform(PUVQFD, verbose = FALSE)

blas_set_num_threads(1)
omp_set_num_threads(1)
options(future.globals.maxSize=50*1024^3) 
E79WZA <- SCTransform(E79WZA, verbose = FALSE)

blas_set_num_threads(1)
omp_set_num_threads(1)
options(future.globals.maxSize=50*1024^3) 
P6KLSUS <- SCTransform(P6KLSUS, verbose = FALSE)

```


```{r}
PUVQFD <- RunPCA(PUVQFD, npcs = 50)
ElbowPlot(PUVQFD, ndims = 50)

E79WZA <- RunPCA(E79WZA, npcs = 50)
ElbowPlot(E79WZA, ndims = 50)

P6KLSUS <- RunPCA(P6KLSUS, npcs = 50)
ElbowPlot(P6KLSUS, ndims = 50)



PUVQFD <- FindNeighbors(PUVQFD, reduction = "pca", dims = 1:30)
PUVQFD <- FindClusters(PUVQFD, resolution = 0.5)
PUVQFD <- RunUMAP(PUVQFD, dims = 1:30)

E79WZA <- FindNeighbors(E79WZA, reduction = "pca", dims = 1:30)
E79WZA <- FindClusters(E79WZA, resolution = 0.5)
E79WZA <- RunUMAP(E79WZA, dims = 1:30)

P6KLSUS <- FindNeighbors(P6KLSUS, reduction = "pca", dims = 1:30)
P6KLSUS <- FindClusters(P6KLSUS, resolution = 0.5)
P6KLSUS <- RunUMAP(P6KLSUS, dims = 1:30)

```


### Plot by processing (fresh/frozen)
```{r fig.width=12, fig.height=4}
p1 <- DimPlot(object = PUVQFD, reduction = 'umap', pt.size = 0.002, label = F, group.by = "processing") +
 theme(legend.position = "none") +
  blank

p2 <- DimPlot(object = E79WZA, reduction = 'umap', pt.size = 0.002, label = F, group.by = "processing") +
 theme(legend.position = "none")+
  blank

p3 <- DimPlot(object = P6KLSUS, reduction = 'umap', pt.size = 0.002, label = F, group.by = "processing") +
 theme(legend.position = "none")+
  blank

p4 <- p1 | p2 | p3
p4
```


#### PLot by major celltype
```{r fig.width=12, fig.height=4}
p1 <- DimPlot(object = PUVQFD, reduction = 'umap', pt.size = 0.002, label = F, group.by = "major_celltype",
              cols = c("#0072B2", "Firebrick", "#009E73","#56B4E9","#E69F00", "#F0E442")) +
 theme(legend.position = "none") +
  blank

p2 <- DimPlot(object = E79WZA, reduction = 'umap', pt.size = 0.002, label = F, group.by = "major_celltype",
              cols = c("#0072B2", "Firebrick", "#009E73","#56B4E9","#E69F00", "#F0E442")) +
 theme(legend.position = "none")+
  blank

p3 <- DimPlot(object = P6KLSUS, reduction = 'umap', pt.size = 0.002, label = F, group.by = "major_celltype",
              cols = c("#0072B2", "Firebrick", "#009E73","#56B4E9","#E69F00", "#F0E442")) +
 theme(legend.position = "none")+
  blank

p4 <- p1 | p2 | p3
p4
```




```{r fig.width=5, fig.height=3}
p1 <- ggplot(PUVQFD@meta.data, aes(x=processing, fill=major_celltype)) + geom_bar(position="fill") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_manual(values=c("#0072B2", "Firebrick", "#009E73","#56B4E9","#E69F00", "#F0E442"))+
 theme(legend.position = "none", panel.background=element_blank(),axis.title.y=element_blank(),axis.title.x=element_blank())

p2 <- ggplot(E79WZA@meta.data, aes(x=processing, fill=major_celltype)) + geom_bar(position="fill") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_manual(values=c("#0072B2", "Firebrick", "#009E73","#56B4E9","#E69F00", "#F0E442"))+
 theme(legend.position = "none", panel.background=element_blank(),axis.title.y=element_blank(),axis.title.x=element_blank())

p3 <- ggplot(P6KLSUS@meta.data, aes(x=processing, fill=major_celltype)) + geom_bar(position="fill") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_manual(values=c("#0072B2", "Firebrick", "#009E73","#56B4E9","#E69F00", "#F0E442"))+
 theme( panel.background=element_blank(),axis.title.y=element_blank(),axis.title.x=element_blank())


p4 <- p1 | p2 | p3
p4
```



### Plot correlation of average gene expression levels in PCs of RRMM12
```{r fig.width=4, fig.height=4}
library(ggplot2)
library(cowplot)
theme_set(theme_cowplot())
Idents(PUVQFD) <- "major_celltype"
PCs <- subset(PUVQFD, idents = "PCs")
Idents(PCs) <- "processing"
avg.PCs <- log1p(AverageExpression(PCs, verbose = FALSE)$SCT)

p1 <- ggplot(avg.PCs, aes(fresh, frozen)) + 
  geom_point() + 
  ggtitle("PCs RRMM12") +
  stat_cor(method = "pearson")
p1

```



