---
title: "K43R_Fig_7 + Supplementary Fig. 7"
output: html_notebook
---


```{r setup}
library(dplyr)
library(tidyverse)
library(Seurat)
library(ggplot2)
library(sctransform)
library(qs)
library(RhpcBLASctl)
library(ggpointdensity)
library(viridis)
library(patchwork)
library(data.table)
library(presto)
library(RColorBrewer)
library(ComplexHeatmap)
library(hypeR)
```

### for reduced UMAP visualization
```{r}
blank <- theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank())
```


### Load complete RRMM dataset (all patients, all time-points, BME + PCs) + metadata
```{r}
K43R_all <- qread(".../K43R_all.qs")


#subset BME
Idents(K43R_all_5) <- "major_population"
K43R_ME_6 <- subset(K43R_all_5, idents = "TME")

```


### check abundance differences per patient
```{r}
df <- FetchData(K43R_ME_6, vars = c("celltype_1", "patient"))
df <- df %>% modify_if(is.character, as.factor)

df_freq <- df %>% 
  group_by(patient) %>% 
  dplyr::count(celltype_1, patient, .drop = F) %>% 
  mutate(prop=prop.table(n)) %>% 
  mutate(prop=format(prop,digits=2))
df_freq$prop <- as.numeric(as.character(df_freq$prop))


# add 1q info here: high = subclonal + dominant; low = ND/rare
df_freq$class_1q <- ifelse(df_freq$patient=="2FTMUU","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="32ARQL","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="37A66E","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="6K8YTY","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="6KLSUS","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="8YGUU8","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="CSU6UB","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="E79WZA","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="N5CC3E","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="N7ZK3L","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="NUL141","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="PUVQFD","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="R1H4HE","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="R1V8RN","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="TS6ZX9","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="VP6CSY","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="XF2DTB","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="YK554L","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="ZPMZFJ","low",as.character(df_freq$class_1q)) 
df_freq$class_1q <- ifelse(df_freq$patient=="ZVNGET","high",as.character(df_freq$class_1q))

```

### check abundance change per celltype
```{r fig.width=14, fig.height=15}

require(ggpubr)
p4 <- ggplot(df_freq, aes(x=factor(class_1q, levels = c("low" , "high")), y=prop)) + 
  geom_boxplot(outlier.size = 0.5) +
  facet_wrap(vars(celltype_1), scales = "free_y", ncol = 6) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.6) +
  geom_text(aes(label=patient),hjust=0, vjust=0, size = 2)+
  theme(strip.text.x = element_text(size = 6)) +
  theme(axis.text.x = element_text(size=7))+
  theme(axis.text.y = element_text(size=6)) +
  theme_bw() + stat_compare_means()
p4

```


### Summarize p-values in barplot
```{r fig.width=3, fig.height=4}
pval <- c(0.77, 0.9, 0.64, 0.43, 1, 0.9, 1, 0.77, 0.97, 0.83, 0.47, 
          0.38, 0.48, 0.83, 0.77, 0.017, 0.84, 0.43, 0.11, 0.3,0.28, 1,0.4,  0.26,
          0.59, 0.3, 0.71, 0.97, 0.84, 0.43, 0.17, 0.48)
celltype_1 <- as.data.frame(table(K43R_ME_6$celltype_1))
celltype_1 <- as.data.frame(celltype_1[,1])
colnames(celltype_1) <- "celltype_1"



celltype_1$pval <- pval 
celltype_1$pval_log <- log10(celltype_1$pval)

p1 <- ggplot(celltype_1, aes(y=pval_log, x=reorder(celltype_1, pval_log))) + 
    geom_bar(position="dodge", stat="identity", color="black") +
 theme(axis.text.x = element_text(angle = 90, hjust = 1), panel.background=element_blank()) +
 geom_hline(yintercept=-1.301, linetype="dashed", color = "black", size = 0.25)+
 ylab("Log(10)P") + xlab("") +coord_flip()

p1
```


### Plot only NK_dim
```{r fig.width=3, fig.height=4}
df_freq2<- df_freq %>% filter (celltype_1=="NK_dim")

require(ggpubr)
p4 <- ggplot(df_freq2, aes(x=factor(class_1q, levels = c("low" ,"high")), y=prop)) + 
  geom_boxplot(outlier.size = 0.5, aes(fill=class_1q)) +
  scale_fill_brewer(palette="Dark2")+
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.6) +
  theme(strip.text.x = element_text(size = 6)) +
  theme(axis.text.x = element_text(size=7))+
  theme(axis.text.y = element_text(size=6)) +
  theme_bw() + stat_compare_means()
p4 
```



### check abundance differences per patient in Macrophages (celltype_2)
```{r}

df <- FetchData(K43R_ME_6, vars = c("celltype_2", "patient"))
df <- df %>% modify_if(is.character, as.factor)

df_freq <- df %>% 
  group_by(patient) %>% 
  dplyr::count(celltype_2, patient, .drop = F) %>% 
  mutate(prop=prop.table(n)) %>% 
  mutate(prop=format(prop,digits=2))
df_freq$prop <- as.numeric(as.character(df_freq$prop))

# add 1q info here: high = subclonal + dominant; low = ND/rare
df_freq$class_1q <- ifelse(df_freq$patient=="2FTMUU","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="32ARQL","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="37A66E","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="6K8YTY","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="6KLSUS","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="8YGUU8","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="CSU6UB","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="E79WZA","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="N5CC3E","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="N7ZK3L","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="NUL141","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="PUVQFD","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="R1H4HE","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="R1V8RN","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="TS6ZX9","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="VP6CSY","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="XF2DTB","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="YK554L","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="ZPMZFJ","low",as.character(df_freq$class_1q)) 
df_freq$class_1q <- ifelse(df_freq$patient=="ZVNGET","high",as.character(df_freq$class_1q))

```


```{r fig.width=14, fig.height=5}

require(ggpubr)
p4 <- ggplot(df_freq, aes(x=factor(class_1q, levels = c("low" , "high")), y=prop)) + 
  geom_boxplot(outlier.size = 0.5) +
  facet_wrap(vars(celltype_2), scales = "free_y", ncol = 6) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.6) +
  geom_text(aes(label=patient),hjust=0, vjust=0, size = 2)+
  theme(strip.text.x = element_text(size = 6)) +
  theme(axis.text.x = element_text(size=7))+
  theme(axis.text.y = element_text(size=6)) +
  theme_bw() + stat_compare_means()
p4

```

```{r fig.width=2.5, fig.height=3.5}
pval <- c(0.84,0.42 ,0.65 ,0.64 ,0.58 ,0.4 ,0.1 ,0.59 ,0.43 ,0.013 ,0.77)
celltype_2 <- as.data.frame(table(K43R_ME_6$celltype_2))
celltype_2 <- as.data.frame(celltype_2[,1])
colnames(celltype_2) <- "celltype_2"

celltype_2$pval <- pval 
celltype_2$pval_log <- log10(celltype_2$pval)

# remove old T-HSP annotation
celltype_2 <- celltype_2[-c(7), ]


p1 <- ggplot(celltype_2, aes(y=pval_log, x=reorder(celltype_2, pval_log))) + 
    geom_bar(position="dodge", stat="identity", color="black") +
 theme(axis.text.x = element_text(angle = 90, hjust = 1), panel.background=element_blank()) +
 geom_hline(yintercept=-1.301, linetype="dashed", color = "black", size = 0.25)+
 ylab("Log(10)P") + xlab("") +coord_flip()

p1
```


## plot TAM3 only
```{r fig.width=3, fig.height=4}
df_freq2<- df_freq %>% filter (celltype_2=="TAM3")

require(ggpubr)
p4 <- ggplot(df_freq2, aes(x=factor(class_1q, levels = c("low" ,"high")), y=prop)) + 
  geom_boxplot(outlier.size = 0.5, aes(fill=class_1q)) +
  scale_fill_brewer(palette="Dark2")+
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.6) +
  theme(strip.text.x = element_text(size = 6)) +
  theme(axis.text.x = element_text(size=7))+
  theme(axis.text.y = element_text(size=6)) +
  theme_bw() + stat_compare_means()
p4 
```



### Check NK cells (especially NK_bright) --> e.g. IL1B & IL18 interaction
```{r}
Idents(K43R_ME_6) <- "celltype_1"
NK <- subset(K43R_ME_6, idents = c("NK_bright", "NK_dim", "NKT"))
```



```{r}
blas_set_num_threads(1)
omp_set_num_threads(1)
options(future.globals.maxSize=50*1024^3) 
NK <- SCTransform(NK, verbose = FALSE)  
```


```{r}
NK <- RunPCA(NK, pcs = 50, verbose = FALSE) ## no "split.by" needed)
ElbowPlot(object = NK, ndims = 50) 
```


```{r}
NK <- RunHarmony(NK, group.by.vars = "sample_id", assay.use="SCT", max.iter.harmony = 20, verbose = F)
```


```{r}
NK <- FindNeighbors(NK, reduction = "harmony", dims = 1:5, verbose = F)
NK <- FindClusters(NK, resolution = 0.2, verbose = F)
NK <- RunUMAP(NK, dims = 1:5, reduction = "harmony", verbose = F)
```


```{r fig.width=12, fig.height=9}
Idents(NK) <- "seurat_clusters"
NK <- RenameIdents(NK, `0` = "Mature_NK", `1` = "NKT", `2` = "Activated_NK", `3` = "NK_bright")

df <- as.data.frame(NK@active.ident)
NK <- AddMetaData(NK, col.name = "celltype_2", metadata = df)
```


```{r}
Idents(NK) <- "celltype_2"
levels(NK) <- c("NK_bright", "Activated_NK","NKT", "Mature_NK")
                                           
df <- as.data.frame(NK@active.ident)
NK <- AddMetaData(NK, col.name = "celltype_2", metadata = df)

```


```{r fig.width=4.3, fig.height=4}
p4 <- DimPlot(object = NK, reduction = 'umap', pt.size = 0.5, label = T, group.by = "celltype_2") + blank
p4
```


```{r fig.width=6, fig.height=3}
p1 <- DotPlot(NK, features = c("FGFBP2","SPON2", "FCGR3A",
                         "CD3D","CD3G", "KLRC2", 
                         "KLRB1","CD69", "JUN",
                         "SELL", "KLRC1","NCAM1"),
        group.by = "celltype_2", dot.scale = 6.7) +
  scale_color_gradient( low="blue",high="red") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
p1
```



```{r fig.width=7, fig.height=2}
Idents(NK) <- "class_1q"

# 1q info here: high = subclonal + dominant; low = ND/rare


NK_low <- subset(NK, idents = "low")
emb <- Embeddings(NK_low, reduction = "umap")
emb <- as.data.frame(emb)
p1<- ggplot(data = emb, mapping = aes(x=UMAP_1, y = UMAP_2)) +
  geom_pointdensity() +
  scale_color_viridis() + blank


NK_high <- subset(NK, idents = "high")
emb <- Embeddings(NK_high, reduction = "umap")
emb <- as.data.frame(emb)
p2 <- ggplot(data = emb, mapping = aes(x=UMAP_1, y = UMAP_2)) +
  geom_pointdensity() +
  scale_color_viridis() + blank



p3 <- FeaturePlot(object = NK,  features = c("GZMB"), cols = c("grey", "firebrick"), 
                 pt.size =0.5, min.cutoff = "q50", order = F) + blank


p4 <- FeaturePlot(object = NK,  features = c("PRF1"), cols = c("grey", "firebrick"), 
                 pt.size =0.5, min.cutoff = "q50", order = F) + blank


p5 <- p1 | p2 | p3 | p4

p5

```


### Add celltype_2 info of NK cells to total ME dataset
```{r}

df <- FetchData(NK, vars = c("celltype_2"))

K43R_ME_6 <- AddMetaData(K43R_ME_6, "celltype_2", metadata = df)
table(K43R_ME_6$celltype_2)
```



```{r}
df <- FetchData(K43R_ME_6, vars = c("celltype_2", "patient"))
df <- df %>% modify_if(is.character, as.factor)

df_freq <- df %>% 
  group_by(patient) %>% 
  dplyr::count(celltype_2, patient, .drop = F) %>% 
  mutate(prop=prop.table(n)) %>% 
  mutate(prop=format(prop,digits=2))
df_freq$prop <- as.numeric(as.character(df_freq$prop))

# 1q info here: high = dominant; int = subclonal; low = ND/rare

df_freq$class_1q <- ifelse(df_freq$patient=="2FTMUU","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="32ARQL","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="37A66E","int",as.character(df_freq$class_1q)) #
df_freq$class_1q <- ifelse(df_freq$patient=="6K8YTY","low",as.character(df_freq$class_1q)) 
df_freq$class_1q <- ifelse(df_freq$patient=="6KLSUS","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="8YGUU8","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="CSU6UB","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="E79WZA","int",as.character(df_freq$class_1q)) #
df_freq$class_1q <- ifelse(df_freq$patient=="N5CC3E","int",as.character(df_freq$class_1q)) # 
df_freq$class_1q <- ifelse(df_freq$patient=="N7ZK3L","high",as.character(df_freq$class_1q)) 
df_freq$class_1q <- ifelse(df_freq$patient=="NUL141","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="PUVQFD","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="R1H4HE","int",as.character(df_freq$class_1q)) #
df_freq$class_1q <- ifelse(df_freq$patient=="R1V8RN","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="TS6ZX9","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="VP6CSY","int",as.character(df_freq$class_1q)) #
df_freq$class_1q <- ifelse(df_freq$patient=="XF2DTB","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="YK554L","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="ZPMZFJ","low",as.character(df_freq$class_1q)) 
df_freq$class_1q <- ifelse(df_freq$patient=="ZVNGET","high",as.character(df_freq$class_1q))
```

```{r fig.width=10, fig.height=3.5}
require(ggpubr)

my_comparisons <- list( c("low", "high"), c("low", "int") )


p4 <- ggplot(df_freq, aes(x=factor(class_1q, levels = c("low", "int","high")), y=prop)) + 
  geom_beeswarm(size=2.5, cex = 3) +
  stat_summary(fun="mean",geom="crossbar", mapping=aes(ymin=..y.., ymax=..y..), 
                 width=0.4,size=0.9, position=position_dodge(),show.legend = FALSE, color = "red") +
 stat_compare_means(comparisons = my_comparisons)+
  facet_wrap(vars(celltype_2), scales = "free_y", ncol = 5) +
  theme(panel.background=element_blank(),
        axis.title.x=element_blank(),
        axis.line = element_line(colour = "black")+
    stat_summary(fun="mean",geom="crossbar", mapping=aes(ymin=..y.., ymax=..y..), width=1, position=position_dodge(),show.legend = FALSE)) + 
  ylab(element_blank()) + 
  ggtitle("NK") +
 theme(axis.text.x = element_text(angle = 45, hjust = 1))
p4 


```



Check Dendritic cells
```{r}
Idents(K43R_ME_6) <- "celltype_1"
DCs <- subset(K43R_ME_6, idents = c("cDC2"))
```


```{r}
blas_set_num_threads(1)
omp_set_num_threads(1)
options(future.globals.maxSize=50*1024^3) 
DCs <- SCTransform(DCs, verbose = FALSE)  

```


```{r}
DCs <- RunPCA(DCs, pcs = 50, verbose = FALSE) ## no "split.by" needed)
ElbowPlot(object = DCs, ndims = 50) 
```



```{r}
DCs <- RunHarmony(DCs, group.by.vars = "sample_id", assay.use="SCT", max.iter.harmony = 20, verbose = F)
```


```{r}
DCs <- FindNeighbors(DCs, reduction = "harmony", dims = 1:10)
DCs <- FindClusters(DCs, resolution = 0.1, verbose = F)
DCs <- RunUMAP(DCs, dims = 1:10, reduction = "harmony", verbose = F)
```


```{r fig.width=12, fig.height=9}
Idents(DCs) <- "seurat_clusters"
DCs <- RenameIdents(DCs, `0` = "cDC2_B", `1` = "cDC2_A",`2`= "cDC2_Cycling", `3`="cDC2_C")

df <- as.data.frame(DCs@active.ident)
DCs <- AddMetaData(DCs, col.name = "celltype_2", metadata = df)
```


```{r}
Idents(DCs) <- "celltype_2"
levels(DCs) <- c("cDC2_A", "cDC2_B", "cDC2_C", "cDC2_Cycling")
                                           
df <- as.data.frame(DCs@active.ident)
DCs <- AddMetaData(DCs, col.name = "celltype_2", metadata = df)

```


```{r fig.width=8, fig.height=4}
p4 <- DimPlot(object = DCs, reduction = 'umap', pt.size = 1, label = T, group.by = "celltype_2") + blank

p5 <- DimPlot(object = DCs, reduction = 'umap', pt.size = 1.5, label = F, group.by = "class_1q", cols = "Dark2") + blank


p6 <- p4 | p5
p6
```


```{r}
Idents(DCs) <- "celltype_2"
DEA_DCs <- FindAllMarkers(DCs, only.pos = F, logfc.threshold = 0.1)
DEA_DCs <- DEA_DCs %>% filter(p_val_adj < 0.05) 


```


```{r fig.width=7.1, fig.height=6}
top10 <- DEA_DCs %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
p1 <- DoHeatmap(DCs, features = top10$gene, draw.lines = F, raster = F) + 
  scale_fill_gradient2( low="blue",high="red", mid = "white") 
p1 
```



### prepare heatmap 
```{r}
Idents(DCs) <- "celltype_2"
DC_avg <- AverageExpression(DCs, assays = "SCT", slot = "data", return.seurat = T, verbose = F)

```


```{r fig.width=4, fig.height=5}
p1 <- DoHeatmap(DC_avg, features = c("CCR1", "CCR2", "CD163", "TREM1", "IL17RA","PYCARD", "SERPINA1",
                               "CLEC12A","CLEC4E","VSIR", "CLEC7A",  
                                "CTSD","CTSS", "CTSB"), draw.lines = F, raster = F) + 
  scale_fill_gradient2( low="blue",high="red", mid = "white") 

p1
```



### Check abundance in whole dataset
```{r}
df <- FetchData(DCs, vars = c("celltype_2"))
K43R_ME_6 <- AddMetaData(K43R_ME_6, "celltype_2", metadata = df)
table(K43R_ME_6$celltype_2)
```


```{r}
df <- FetchData(K43R_ME_6, vars = c("celltype_2", "patient"))
df <- df %>% modify_if(is.character, as.factor)

df_freq <- df %>% 
  group_by(patient) %>% 
  dplyr::count(celltype_2, patient, .drop = F) %>% 
  mutate(prop=prop.table(n)) %>% 
  mutate(prop=format(prop,digits=2))
df_freq$prop <- as.numeric(as.character(df_freq$prop))


# 1q info here: high = dominant; int = subclonal; low = ND/rare
df_freq$class_1q <- ifelse(df_freq$patient=="2FTMUU","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="32ARQL","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="37A66E","int",as.character(df_freq$class_1q)) #
df_freq$class_1q <- ifelse(df_freq$patient=="6K8YTY","low",as.character(df_freq$class_1q)) 
df_freq$class_1q <- ifelse(df_freq$patient=="6KLSUS","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="8YGUU8","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="CSU6UB","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="E79WZA","int",as.character(df_freq$class_1q)) #
df_freq$class_1q <- ifelse(df_freq$patient=="N5CC3E","int",as.character(df_freq$class_1q)) # 
df_freq$class_1q <- ifelse(df_freq$patient=="N7ZK3L","high",as.character(df_freq$class_1q)) 
df_freq$class_1q <- ifelse(df_freq$patient=="NUL141","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="PUVQFD","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="R1H4HE","int",as.character(df_freq$class_1q)) #
df_freq$class_1q <- ifelse(df_freq$patient=="R1V8RN","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="TS6ZX9","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="VP6CSY","int",as.character(df_freq$class_1q)) #
df_freq$class_1q <- ifelse(df_freq$patient=="XF2DTB","low",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="YK554L","high",as.character(df_freq$class_1q))
df_freq$class_1q <- ifelse(df_freq$patient=="ZPMZFJ","low",as.character(df_freq$class_1q)) 
df_freq$class_1q <- ifelse(df_freq$patient=="ZVNGET","high",as.character(df_freq$class_1q))
```



```{r fig.width=10, fig.height=4}
require(ggpubr)

my_comparisons <- list( c("low", "high") , c("low", "int") )

p4 <- ggplot(df_freq, aes(x=factor(class_1q, levels = c("low", "int","high")), y=prop)) + 
  geom_beeswarm(size=2.5, cex = 3) +
  stat_summary(fun="mean",geom="crossbar", mapping=aes(ymin=..y.., ymax=..y..), 
                 width=0.4,size=0.9, position=position_dodge(),show.legend = FALSE, color = "red") +
 stat_compare_means(comparisons = my_comparisons)+
  facet_wrap(vars(celltype_2), scales = "free_y", ncol = 5) +
  theme(panel.background=element_blank(),
        axis.title.x=element_blank(),
        axis.line = element_line(colour = "black")+
    stat_summary(fun="mean",geom="crossbar", mapping=aes(ymin=..y.., ymax=..y..), width=1, position=position_dodge(),show.legend = FALSE)) + 
  ylab(element_blank()) + 
  ggtitle("DC") +
 theme(axis.text.x = element_text(angle = 45, hjust = 1))
p4 


```









